|

### 2020-08-20 EvolvertPresent.Rmd
Plots for the group meeting

---
title: "The evolution of badges of status with learners"
author: "Andrés Quiñones"
bibliography: SocialMorphCue.bib
output: 
    beamer_presentation: 
      fig_caption: yes
      keep_tex: yes
header-includes: 
    \usepackage{float}
    \usepackage{multicol}
    \usepackage{amsmath} 
    \usepackage{graphicx}
    \usepackage{array}
                #\floatplacement{figure}{H}
---

```{r loadData, echo=FALSE,include=FALSE,warning=FALSE,message=FALSE}
# Required libraries -----------------------------------------------------------

library(here)
here()
source(here("AccFunc.R"))
# Scenario to be plotted - corresponds to folders where simulations are stored
scenario<-"mutType"
(listTest<-list.files(here("Simulations",paste0(scenario,"_"))))
(evolList<-grep(".txt",grep("evol",listTest,value=TRUE),value=TRUE))
(indList<-grep("ind",listTest,value=TRUE))

fileId<-1
evol<-fread(here("Simulations",paste0(scenario,"_"),evolList[fileId]))
pop<-fread(here("Simulations",paste0(scenario,"_"),indList[fileId]))
evolStats<-evol[,.(m.freqGenHawk=mean(freqGenHawks),
                   upIQR.freqGenHawk=fivenum(freqGenHawks)[4],
                   lowIQR.freqGenHawk=fivenum(freqGenHawks)[2],
                   m.freqGenDove=mean(freqGenDove),
                   upIQR.freqGenDove=fivenum(freqGenDove)[4],
                   lowIQR.freqGenDove=fivenum(freqGenDove)[2],
                   m.freqEval=mean(freqGenEval),
                   upIQR.freqGenEval=fivenum(freqGenEval)[4],
                   lowIQR.freqGenEval=fivenum(freqGenEval)[2], 
                   m.freqFenHawk=mean(freqFenHawks),
                   upIQR.freqFenHawk=fivenum(freqFenHawks)[4],
                   lowIQR.freqFenHawk=fivenum(freqFenHawks)[2],
                   m.freqFenDove=mean(freqFenDoves),
                   upIQR.freqFenDove=fivenum(freqFenDoves)[4],
                   lowIQR.freqFenDove=fivenum(freqFenDoves)[2],
                   m.meanAlpha=mean(meanAlpha),
                   upIQR.alpha=fivenum(meanAlpha)[4],
                   lowIQR.alpha=fivenum(meanAlpha)[2],
                   m.meanBeta=mean(meanBeta),
                   upIQR.beta=fivenum(meanBeta)[4],
                   lowIQR.beta=fivenum(meanBeta)[2],
                   m.freqHH = mean(freqHH),
                   m.freqHD = mean(freqHD),
                   m.freqDD = mean(freqDD),
                   upIQR.freqHH = fivenum(freqHH)[4],
                   upIQR.freqHD = fivenum(freqHD)[4],
                   upIQR.freqDD = fivenum(freqDD)[4],
                   lowIQR.freqHH = fivenum(freqHH)[4],
                   lowIQR.freqHD = fivenum(freqHD)[2],
                   lowIQR.freqDD = fivenum(freqDD)[2],
                   m.weightAct_0=mean(WeightAct_0),
                   m.weightAct_1=mean(WeightAct_1),
                   m.weightAct_2=mean(WeightAct_2),
                   m.weightAct_3=mean(WeightAct_3),
                   m.weightAct_4=mean(WeightAct_4),
                   m.weightCrit_0=mean(WeightCrit_0),
                   m.weightCrit_1=mean(WeightCrit_1),
                   m.weightCrit_2=mean(WeightCrit_2),
                   m.weightCrit_3=mean(WeightCrit_3),
                   m.weightCrit_4=mean(WeightCrit_4)),by=time]
# Extract means and IQR for the dynamic variables ------------------------------

popStats<-pop[, as.list(unlist(lapply(.SD, 
                                      function(x) list(m = mean(x),
                                                       upIQR = fivenum(x)[4],
                                                       downIQR = fivenum(x)[2]
                                      )))),
              by = time, 
              .SDcols=c("Quality","alpha","beta","Badge","nInteract","WeightAct_0",
                        "WeightAct_1","WeightAct_2","WeightAct_3","WeightAct_4",
                        "WeightCrit_0","WeightCrit_1","WeightCrit_2",
                        "WeightCrit_3","WeightCrit_4")]

# knitr::include_graphics("Simulations/mutType_/basicHawkDove.png")
```

@@

```{r,echo=FALSE,fig.align="center",out.width='50%'}
knitr::include_graphics("Images//HDgame.jpeg")
```

@@

```{r, echo=FALSE, fig.cap="\\label{fig:HD_game}Hawk-dove game. Dashed line is the  game theoretical prediction for frequency of hawks.", out.width = '80%'}
knitr::include_graphics("Simulations/mutType_/basicHawkDove.png")
```

@@

```{r fig2, echo=FALSE, fig.cap="\\label{fig:learning_cartoonRBF}Function approximation for the actor and the critic. Using random values for the responses of each center. Circles show the location of the centers and the maximum response they trigger, while the lines show the total response along the badge size continium.", out.width = '60%'}
knitr::include_graphics("cartoonRBF.png")
```

@@

```{r, echo=FALSE,fig.align='center', out.width = '60%'}
knitr::include_graphics("cartoonRBF.png")
```

@@

## When individuals do _NOT_ vary in quality

```{r , echo=FALSE, fig.align='center',out.width = '80%'}
knitr::include_graphics("Simulations//QualStDv_//hawkDoveLearn.png")
```

## When individuals _DO_ vary in quality

```{r, echo=FALSE, fig.align='center', out.width = '80%'}
knitr::include_graphics("Simulations//QualStDv_//hawkDoveLearn_1.png")
```

## Overal effect of quality variation

```{r fig5, echo=FALSE, fig.align='center', out.width = '80%'}
knitr::include_graphics("Simulations//QualStDv_//effectQualVariance.png")
```

## How does variation among _learners_ look like?

```{r fig6, echo=FALSE, fig.align='center',out.width = '90%'}
knitr::include_graphics("Simulations//QualStDv_//WeightsVarQualSt.png")
```

## How do the learning dynamics look like?

```{r fig7, echo=FALSE, out.width = '80%'}
knitr::include_graphics("Simulations//mutType_//learnDyn200.png")
```

## How do learners behave when signals are honest?

```{r fig9, echo=FALSE,out.width='90%',fig.show='hold'}
knitr::include_graphics("Simulations//learHonest_//QualStDv_//weightsVarQualSt.png")
```

## What parameters influence responsiveness?

### Distribution of quality

```{r,echo=FALSE,fig.align='center',out.width='90%'}
knitr::include_graphics("Simulations//learHonest_//QualStDv_//compDistrNorm1Unif.png")
```

## What parameters influence responsiveness?

### Distribution of quality

```{r,echo=FALSE,fig.align='center',out.width='90%'}
knitr::include_graphics("Simulations//learHonest_//QualStDv_//compDistrDiff.png")
```

## What parameters influence responsiveness?

### Strength of quality influence on competition outcome

```{r,echo=FALSE,fig.align='center',out.width='90%'}
knitr::include_graphics("Simulations//learHonest_//strQual_//compStr1_last.png")
```

## What parameters influence responsiveness?

### Strength of quality influence on competition outcome

```{r,echo=FALSE,out.width='45%',fig.show='hold'}
knitr::include_graphics("Simulations//learHonest_//strQual_//compStrDiff.png")
knitr::include_graphics("Simulations//learHonest_//strQual_//compStrpDove.png")
```

# Why the hell is responsiveness so low?

## How do learners fare against pure strategies?

```{r fig8, echo=FALSE, fig.align='center', out.width = '80%'}
knitr::include_graphics("Simulations//mutType_//evolDynTypes.png")
```

## Concluding

* Learning is good, even if you do not learn about your opponent. You still learn about you
* Learning reduces aggressive interactions
* Learning splits up individuals into _types_

## What's next?

* Let reaction norm evolve, under different initial conditions.
* Let learning parmeters evolve
* Let the communication system co-evolve

>>>>

---
title: "Badges of status and IR mediated by learning"
author: "Andrés Quiñones"
bibliography: SocialMorphCue.bib
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
    beamer_presentation: 
      fig_caption: yes
      keep_tex: yes
header-includes: 
    \usepackage{float}
    \usepackage{multicol}
    \usepackage{amsmath} 
    \usepackage{graphicx}
    \usepackage{array}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("../R_files/posPlots.R")
require("here")
source(here("AccFunc.R"))

@@

```{r,echo=FALSE,fig.align="center",out.width='50%'}
knitr::include_graphics("Images//HDgame.png")
```

@@

<!-- ## The hawk-dove game -->

<!-- ```{r, echo=FALSE, fig.cap="\\label{fig:HD_game}Hawk-dove game. Dashed line is the  game theoretical prediction for frequency of hawks.", out.width = '80%'} -->
<!-- knitr::include_graphics("Simulations/mutType_/basicHawkDove.png") -->
<!-- ``` -->

@@

```{r fig2, echo=FALSE, fig.cap="\\label{fig:learning_cartoonRBF}Function approximation for the actor and the critic. Using random values for the responses of each center. Circles show the location of the centers and the maximum response they trigger, while the lines show the total response along the badge size continium.", out.width = '60%'}
knitr::include_graphics("Images/cartoonRBF.png")
```

@@

```{r, echo=FALSE,fig.align='center', out.width = '60%'}
knitr::include_graphics("Images/cartoonRBF.png")
```

@@

## Learning under honest signalling

```{r, echo=FALSE,fig.align='center', out.width = '100%'}
knitr::include_graphics("Simulations/alphaAct_/alphaAct0.6learnDyn.png")
```

## Learning can mediate the evolution of Badges of Status
### Variation among replicates
```{r, echo=FALSE,fig.align='center', out.width='100%'}
knitr::include_graphics("Simulations/betCostEvol1_/evolDynAll_betCost5.png")
```

## Learning can mediate the evolution of Badges of Status
### Individual replicates: Badges
```{r, echo=FALSE,fig.align='center', out.width='100%'}
knitr::include_graphics("Simulations/betCostEvol1_/evolDyn0_betCost5.png")
```

## Learning can mediate the evolution of Badges of Status
### Individual replicates: no badges
```{r, echo=FALSE,fig.align='center', out.width='100%'}
knitr::include_graphics("Simulations/betCostEvol1_/evolDyn5_betCost5.png")
```

## The effect of costs on Badges of Status
### Behavioural interactions

```{r, echo=FALSE,fig.align='center', out.width='100%'}
knitr::include_graphics("Simulations/betCostEvol1_/BehavIntALL.png")
```

### 2020-08-20 EvolvertPresent.tex
Plots for the group meeting

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  ignorenonframetext,
]{beamer}
\usepackage{pgfpages}
\setbeamertemplate{caption}[numbered]
\setbeamertemplate{caption label separator}{: }
\setbeamercolor{caption name}{fg=normal text.fg}
\beamertemplatenavigationsymbolsempty
% Prevent slide breaks in the middle of a paragraph:
\widowpenalties 1 10000
\raggedbottom
\setbeamertemplate{part page}{
  \centering
  \begin{beamercolorbox}[sep=16pt,center]{part title}
    \usebeamerfont{part title}\insertpart\par
  \end{beamercolorbox}
}
\setbeamertemplate{section page}{
  \centering
  \begin{beamercolorbox}[sep=12pt,center]{part title}
    \usebeamerfont{section title}\insertsection\par
  \end{beamercolorbox}
}
\setbeamertemplate{subsection page}{
  \centering
  \begin{beamercolorbox}[sep=8pt,center]{part title}
    \usebeamerfont{subsection title}\insertsubsection\par
  \end{beamercolorbox}
}
\AtBeginPart{
  \frame{\partpage}
}
\AtBeginSection{
  \ifbibliography
  \else
    \frame{\sectionpage}
  \fi
}
\AtBeginSubsection{
  \frame{\subsectionpage}
}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={The evolution of badges of status with learners},
  pdfauthor={Andrés Quiñones},
  pdfborder={0 0 0},
  breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\newif\ifbibliography
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-2}

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{float} \usepackage{multicol} \usepackage{amsmath} \usepackage{graphicx} \usepackage{array}

\title{The evolution of badges of status with learners}
\author{Andrés Quiñones}
\date{}

@@

\begin{frame}{The hawk-dove game}
\protect\hypertarget{the-hawk-dove-game-2}{}

\begin{figure}
\includegraphics[width=0.8\linewidth]{Simulations/mutType_/basicHawkDove} \caption{\label{fig:HD_game}Hawk-dove game. Dashed line is the  game theoretical prediction for frequency of hawks.}\label{fig:unnamed-chunk-3}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.6\linewidth]{cartoonRBF} \caption{\label{fig:learning_cartoonRBF}Function approximation for the actor and the critic. Using random values for the responses of each center. Circles show the location of the centers and the maximum response they trigger, while the lines show the total response along the badge size continium.}\label{fig:fig2}
\end{figure}

@@

\begin{center}\includegraphics[width=0.5\linewidth]{QD_presentation_files/figure-beamer/unnamed-chunk-9-1} \end{center}

@@

\begin{center}\includegraphics[width=0.5\linewidth]{QD_presentation_files/figure-beamer/unnamed-chunk-10-1} \end{center}

@@

\begin{center}\includegraphics[width=0.6\linewidth]{cartoonRBF} \end{center}

@@

\begin{frame}{When individuals do \emph{NOT} vary in quality}
\protect\hypertarget{when-individuals-do-not-vary-in-quality}{}

\begin{center}\includegraphics[width=0.8\linewidth]{Simulations//QualStDv_//hawkDoveLearn} \end{center}

@@

\begin{frame}{When individuals \emph{DO} vary in quality}
\protect\hypertarget{when-individuals-do-vary-in-quality}{}

\begin{center}\includegraphics[width=0.8\linewidth]{Simulations//QualStDv_//hawkDoveLearn_1} \end{center}

\begin{frame}{Overal effect of quality variation}
\protect\hypertarget{overal-effect-of-quality-variation}{}

\begin{center}\includegraphics[width=0.8\linewidth]{Simulations//QualStDv_//effectQualVariance} \end{center}

\begin{frame}{How does variation among \emph{learners} look like?}
\protect\hypertarget{how-does-variation-among-learners-look-like}{}

\begin{center}\includegraphics[width=0.9\linewidth]{Simulations//QualStDv_//WeightsVarQualSt} \end{center}

\begin{frame}{How do the learning dynamics look like?}
\protect\hypertarget{how-do-the-learning-dynamics-look-like}{}

\includegraphics[width=0.8\linewidth]{Simulations//mutType_//learnDyn200}

\begin{frame}{How do learners behave when signals are honest?}
\protect\hypertarget{how-do-learners-behave-when-signals-are-honest}{}

\includegraphics[width=0.9\linewidth]{Simulations//learHonest_//QualStDv_//weightsVarQualSt}

\begin{frame}{What parameters influence responsiveness?}
\protect\hypertarget{what-parameters-influence-responsiveness}{}

\begin{block}{Distribution of quality}

\begin{center}\includegraphics[width=0.9\linewidth]{Simulations//learHonest_//QualStDv_//compDistrNorm1Unif} \end{center}

@@

\begin{frame}{What parameters influence responsiveness?}
\protect\hypertarget{what-parameters-influence-responsiveness-1}{}

\begin{block}{Distribution of quality}

\begin{center}\includegraphics[width=0.9\linewidth]{Simulations//learHonest_//QualStDv_//compDistrDiff} \end{center}

@@

\begin{frame}{What parameters influence responsiveness?}
\protect\hypertarget{what-parameters-influence-responsiveness-2}{}

\begin{block}{Strength of quality influence on competition outcome}

\begin{center}\includegraphics[width=0.9\linewidth]{Simulations//learHonest_//strQual_//compStr1_last} \end{center}

@@

\begin{frame}{What parameters influence responsiveness?}
\protect\hypertarget{what-parameters-influence-responsiveness-3}{}

\begin{block}{Strength of quality influence on competition outcome}

\includegraphics[width=0.45\linewidth]{Simulations//learHonest_//strQual_//compStrDiff}
\includegraphics[width=0.45\linewidth]{Simulations//learHonest_//strQual_//compStrpDove}

@@

\hypertarget{why-the-hell-is-responsiveness-so-low}{%
\section{Why the hell is responsiveness so
low?}\label{why-the-hell-is-responsiveness-so-low}}

\begin{frame}{How do learners fare against pure strategies?}
\protect\hypertarget{how-do-learners-fare-against-pure-strategies}{}

\begin{center}\includegraphics[width=0.8\linewidth]{Simulations//mutType_//evolDynTypes} \end{center}

\begin{frame}{Concluding}
\protect\hypertarget{concluding}{}

\begin{itemize}
\tightlist
\item
  Learning is good, even if you do not learn about your opponent. You
  still learn about you
\item
  Learning reduces aggressive interactions
\item
  Learning splits up individuals into \emph{types}
\end{itemize}

\begin{frame}{What's next?}
\protect\hypertarget{whats-next}{}

\begin{itemize}
\tightlist
\item
  Let reaction norm evolve, under different initial conditions.
\item
  Let learning parmeters evolve
\item
  Let the communication system co-evolve
\end{itemize}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  ignorenonframetext,
]{beamer}
\usepackage{pgfpages}
\setbeamertemplate{caption}[numbered]
\setbeamertemplate{caption label separator}{: }
\setbeamercolor{caption name}{fg=normal text.fg}
\beamertemplatenavigationsymbolsempty
% Prevent slide breaks in the middle of a paragraph
\widowpenalties 1 10000
\raggedbottom
\setbeamertemplate{part page}{
  \centering
  \begin{beamercolorbox}[sep=16pt,center]{part title}
    \usebeamerfont{part title}\insertpart\par
  \end{beamercolorbox}
}
\setbeamertemplate{section page}{
  \centering
  \begin{beamercolorbox}[sep=12pt,center]{part title}
    \usebeamerfont{section title}\insertsection\par
  \end{beamercolorbox}
}
\setbeamertemplate{subsection page}{
  \centering
  \begin{beamercolorbox}[sep=8pt,center]{part title}
    \usebeamerfont{subsection title}\insertsubsection\par
  \end{beamercolorbox}
}
\AtBeginPart{
  \frame{\partpage}
}
\AtBeginSection{
  \ifbibliography
  \else
    \frame{\sectionpage}
  \fi
}
\AtBeginSubsection{
  \frame{\subsectionpage}
}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Badges of status and IR mediated by learning},
  pdfauthor={Andrés Quiñones},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\newif\ifbibliography
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{float} \usepackage{multicol} \usepackage{amsmath} \usepackage{graphicx} \usepackage{array}

\title{Badges of status and IR mediated by learning}
\author{Andrés Quiñones}
\date{August 20, 2020}

@@

@@

\begin{figure}
\includegraphics[width=0.6\linewidth]{Images/cartoonRBF} \caption{\label{fig:learning_cartoonRBF}Function approximation for the actor and the critic. Using random values for the responses of each center. Circles show the location of the centers and the maximum response they trigger, while the lines show the total response along the badge size continium.}\label{fig:fig2}
\end{figure}

@@

\begin{center}\includegraphics[width=0.5\linewidth]{EvolvertPresent_files/figure-beamer/unnamed-chunk-8-1} \end{center}

@@

\begin{center}\includegraphics[width=0.5\linewidth]{EvolvertPresent_files/figure-beamer/unnamed-chunk-9-1} \end{center}

@@

\begin{center}\includegraphics[width=0.6\linewidth]{Images/cartoonRBF} \end{center}

@@

\begin{frame}{Learning under honest signalling}
\protect\hypertarget{learning-under-honest-signalling}{}

\begin{center}\includegraphics[width=1\linewidth]{Simulations/alphaAct_/alphaAct0.6learnDyn} \end{center}

@@

\begin{frame}{Learning can mediate the evolution of Badges of Status}
\protect\hypertarget{learning-can-mediate-the-evolution-of-badges-of-status}{}

\begin{block}{Variation among replicates}

\begin{center}\includegraphics[width=1\linewidth]{Simulations/betCostEvol1_/evolDynAll_betCost5} \end{center}

@@

\begin{frame}{Learning can mediate the evolution of Badges of Status}
\protect\hypertarget{learning-can-mediate-the-evolution-of-badges-of-status-1}{}

\begin{block}{Individual replicates: Badges}

\begin{center}\includegraphics[width=1\linewidth]{Simulations/betCostEvol1_/evolDyn0_betCost5} \end{center}

@@

\begin{frame}{Learning can mediate the evolution of Badges of Status}
\protect\hypertarget{learning-can-mediate-the-evolution-of-badges-of-status-2}{}

\begin{block}{Individual replicates: no badges}

\begin{center}\includegraphics[width=1\linewidth]{Simulations/betCostEvol1_/evolDyn5_betCost5} \end{center}

@@

\begin{frame}{The effect of costs on Badges of Status}
\protect\hypertarget{the-effect-of-costs-on-badges-of-status}{}

\begin{block}{Behavioural interactions}

\begin{center}\includegraphics[width=1\linewidth]{Simulations/betCostEvol1_/BehavIntALL} \end{center}

@@

### 2020-08-20 manuscript_1.0.Rmd
Plots for the group meeting

---
title: |
  | Learning can mediate both badges of status and individual recognition
author:
  - Andrés E. Quiñones^[Laboratorio de Biología Evolutiva de Vertebrados, Departamento de Ciencias Biológicas, Universidad de los Andes, Bogotá, Colombia]
  - C. Daniel Cadena^[Laboratorio de Biología Evolutiva de Vertebrados, Departamento de Ciencias Biológicas, Universidad de los Andes, Bogotá, Colombia]
  - Olof Leimar^[Department of Zoology, Stockholm University, Stockholm, Sweden]
  - Redouan Bshary^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland]
date: "`r format(Sys.time(), '%B %d, %Y')`"
bibliography: SocialMorphCue.bib
style: C:\Users\andre\Zotero\styles\elife.csl
header-includes: 
  - \usepackage{float}
  - \usepackage{setspace}\doublespacing
  - \linespread{2}
  - \usepackage{lineno}
  - \linenumbers
spacing: double
abstract: My abstract which has to be long enough to take multiple
  lines otherwise one does not see the effect of single-spacing.
keywords: "My keywords"
fig_caption: yes
keep_tex: yes
fontsize: 12pt
linespread: 2
output: pdf_document
---

@@

For social animals the outcome of interactions with con-specifics is an important 
determinant of fitness. Irrespective of whether such interactions are of cooperative 
or competitive nature, individual's actions and their partners will partly determine 
their success in the reproductive arena. However, the best action for an individual 
might vary depending on it's own conditions, and with whom they interact. 
Thus, an individual making their action dependent on information about interacting 
partners is typically adaptive [@quinones_Negotiation_2016].
The source of this information however is far from trivial. 
In some cases interacting partners (signalers) might be 'willing' to provide accurate
information, but in others it might be in their own interest to conceal information
[@johnstone_Recognition_1997], or to provide erroneous information [@johnstone_Badges_1993]. 
Typically, in a given context, some proportion of individuals will be interested 
in broadcasting accurate information, and some others in hiding it. 
Take for instance an interaction between two individuals where one can help the other. 
Given some conditions of cost and benefits, the donor will be interested in helping 
a related individual. So for a relative of the donor, 
broadcasting kinship is advantageous; while for an unrelated individual, 
concealing the lack of kinship will be better. This same pattern applies to 
many aspects of social life such as finding mates, feeding offspring, dominance relationships 
and aggressive interactions. In dominance and aggressive interactions an important 
information to make actions dependent on is interacting partners fighting ability, 
which is sometimes referred to as Resource Holding Potential (RHP) [@parker_Assessment_1974], 
or as we will here simply quality. This responsiveness to the quality of the partner is achieved in
communication systems such as badges of status (BS), where an arbitrary signal conveys quality. 
BS can be evolutionarily stable whenever the signal imposes a fitness cost which 
decreases with the quality of an individual  [@botero_Evolution_2010; @johnstone_Badges_1993]. 
So, low quality individuals can't afford the cost of the signal, 
and it's no longer in their interest to fake quality. 
The cost of the honest signal is a more general principle of communications systems, 
usually referred to as the handicap principle [@grafen_Biological_1990; @zahavi_Mate_1975]. 
An alternative system of communication is one where individuals gather and store information
about each individual they interact with. This information, directly collected by 
each individual, should reliably allow them to react adaptively to the 
quality difference with their opponent. This type of assessment strategy is often referred to as 
individual recognition [IR, @whitfield_Plumage_1986]. Evidently, IR is restricted to 
certain interaction structures, individuals are limited by cognitive adaptations in how many
individuals they can recognize and track. Thus, IR and BS have been assumed to be alternative
communication systems fulfilling the same function, but expected to evolve in populations 
under different interaction structures [@dale_Signaling_2001; @sheehan_There_2016].

>>>>

---
title: |
  | Learning can mediate both badges of status and individual recognition
author:
- Andrés E. Quiñones^[Laboratorio de Biología Evolutiva de Vertebrados, Departamento
  de Ciencias Biológicas, Universidad de los Andes, Bogotá, Colombia]
- C. Daniel Cadena*
- Olof Leimar^[Department of Zoology, Stockholm University, Stockholm, Sweden]
- Redouan Bshary^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland]
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  word_document: default
  pdf_document: default
fig_caption: yes
fontsize: 12pt
header-includes:
- \usepackage{float}
- \usepackage{setspace}\doublespacing
- \linespread{2}
- \usepackage{lineno}
- \linenumbers
keep_tex: yes
keywords: My keywords
linespread: 2
bibliography: SocialMorphCue.bib
spacing: double
style: C:\Users\andre\Zotero\styles\elife.csl
abstract: Learning rules.
---

@@

For social animals the outcome of interactions with con-specifics is an important 
determinant of fitness. Irrespective of whether such interactions are of cooperative 
or competitive nature, individual's actions and their partners will partly determine 
their success in the reproductive arena. However, the best action for an individual 
might vary depending on it's own conditions, and with whom they interact. 
Thus, an individual making their action dependent on information about interacting 
partners is typically adaptive [@quinones_Negotiation_2016].
The source of this information however is far from trivial. 
In some cases interacting partners (signalers) might be 'willing' to provide accurate
information, but in others it might be in their own interest to conceal information
[@johnstone_Recognition_1997], or to provide erroneous information [@johnstone_Badges_1993]. 
Typically, in a given context, some proportion of individuals will be interested 
in broadcasting accurate information, and some others in hiding it. 
Take for instance an interaction between two individuals where one can help the other. 
Given some conditions of cost and benefits, the donor will be interested in helping 
a related individual. So for a relative of the donor, 
broadcasting kinship is advantageous; while for an unrelated individual, 
concealing the lack of kinship will be better. This same pattern applies to 
many aspects of social life such as finding mates, feeding offspring, dominance relationships 
and aggressive interactions. In dominance and aggressive interactions an important 
information to make actions dependent on is interacting partners fighting ability, 
which is sometimes referred to as Resource Holding Potential (RHP) [@parker_Assessment_1974], 
or as we will here simply quality. This responsiveness to the quality of the 
partner is achieved in communication systems such as badges of status (BS), 
where an arbitrary signal conveys quality. 
BS can be evolutionarily stable whenever the signal imposes a fitness cost which 
decreases with the quality of an individual  [@botero_Evolution_2010; @johnstone_Badges_1993]. 
So, low quality individuals can't afford the cost of the signal, 
and it's no longer in their interest to fake quality. 
The cost of the honest signal is a more general principle of communications systems, 
usually referred to as the handicap principle [@grafen_Biological_1990; @zahavi_Mate_1975]. 
An alternative system of communication is one where individuals gather and store information
about each individual they interact with. This information, directly collected by 
each individual, should reliably allow them to react adaptively to the 
quality difference with their opponent. This type of assessment strategy is often 
referred to as individual recognition [IR, @whitfield_Plumage_1986]. Evidently, 
IR is restricted to certain interaction structures, individuals are limited by cognitive 
abilities in how many individuals they can recognize and track. Thus, IR and BS 
have been assumed to be alternative communication systems fulfilling the same function, 
but expected to evolve in populations under different conditions
[@dale_Signaling_2001; @sheehan_There_2016]. There is no clear cut reason, however, 
why both systems could not work together, in fact there is some evidence
that they can [@chaine_Manipulating_2018].

A somewhat ignored aspect of communication systems, in the context of aggressive 
interactions, is the cognitive aspects of the receiver module. 
BS are often assume to be cognitively less demanding than IR 
[@sheehan_There_2016]. The idea is that using IR, individuals need to store identity 
and associated information about each individual they interact with. All this 
information requires more memory and therefore is more demanding. In contrast,
the response using BS have been thought as an innate response 
[@botero_Evolution_2010; @johnstone_Badges_1993], where individuals use their 
own quality and the opponent's badge to determine whether to aggressively contest 
a resource. An alternative view of BS is that individuals learn to react to the signal
based on their experiences [@guilford_Receiver_1991]. In line with that view, both
communication systems would require the receivers to learn an association between
the signal and the fighting ability of the bearer. In BS, this association would be a 
a simple monotonic increase of fighting ability with the size of the badge, 
and would be reinforced by every interaction, so in  principle it would be 
quicker to learn. While in IR, the association would 
be different for every individual, depending on the interacting partners. 
Learning could be a unifying cognitive mechanism behind both types of communication 
systems. This idea of learning being a general response mechanism has not been 
thoroughly explored, nor in the empirical or theoretical literature.

Associative learning is a key cognitive mechanism that allows individuals to associate
rewards with environmental stimuli and appropriate behavior [@staddon_Adaptive_2016]. 
Since it was originally described, associative learning has been found in all major
vertebrate taxon, and many invertebrates [@heyes_Simple_2012;
@macphail_Brain_1982; @staddon_Adaptive_2016; @behrens_Associative_2008].
Using these associations to respond to environmental changes is particularly adaptive in
complex environments [@dridi_Environmental_2016]. Besides its wide taxonomic and
ecological relevance, associative learning has shown to be a flexible cognitive mechanism
that can be tuned to different roles depending on the ecological situation 
[@enquist_Power_2016; @quinones_Reinforcement_2019]. Despite all this, associative 
learning is not often included in evolutionary explanations 
[@fawcett_Exposing_2013; @kamil_Optimal_1983; @mcauliffe_Psychology_2015]. 
Computational models of evolution can play an important role to overcome the 
lack of integration between learning and evolution. Reinforcement learning theory 
encompases a series of computational methods inspired on the psychological and 
neurological mechanisms of associative learning [@sutton_Reinforcement_2018].
These set of algorithms allow the implementation of biologically realistic problems, 
capturing the essence of learning processes [@frankenhuis_Enriching_2018; @quinones_Reinforcement_2019]. 
Furthermore, these algorithms can be embebed in evolutionary simulations in order 
to have theoretical predictions of the effect that learning can have in behavioral [@leimar_Learning_2019]
and morphological evolution.

In here we present an evolutionary model where individuals use associative learning 
to develop a tendency to behave aggressively (or  peacefully), in the context of
competition over resoureces, depending on the quantitative morphological trait 
they perceive in their opponent (badge). Over evolutionary time individuals evolve
the size of their badge, and whether it depends on their quality. Under this simple
set up, individuals can use the badge both as a signal of quality (BS)
or as a trait for IR. We use the model to asses under what conditions of interaction
structure we expect different communication signals evolve.
